---
title: "Data Harvesting - Spanish Newspapers"
format: html
editor: visual
---

# Overview

This document scrapes the international sections of four major Spanish newspapers: **El Diario**, **El Mundo**, **El País**, and **La Vanguardia**. For each newspaper we collect article titles, authors, URLs, dates, city of writing origin, countries mentioned, political figures mentioned, and number of comments. At the end, all four datasets are merged into a single CSV with a `newspaper` column and an `article_id` column.

------------------------------------------------------------------------

# 1. El Diario

## Setup

```{r}
library(xml2)
library(rvest)
library(httr)
library(dplyr)
library(stringr)
library(purrr)
library(countrycode)

httr::set_config(
  httr::user_agent("Mozilla/5.0 (Windows NT 10.0; Win64; x64); Carlos Morán / carlos.morangq@gmail.com")
)
```

## Download HTML

```{r}
# Download the HTML of El Diario's International section
eldiario_url <- read_html("https://www.eldiario.es/internacional/")
```

## Titles and URLs (International filter)

The front page of El Diario's international section mixes articles from other sections (society, culture, etc.). We keep only articles whose URL contains `/internacional/`.

```{r}
# Extract titles and URLs from all articles on the page
# The title and URL are in the same <a> inside <h2 class="ni-title">
titles_all <- eldiario_url |>
  html_elements("h2.ni-title a") |>
  html_text2()

ref_all <- eldiario_url |>
  html_elements("h2.ni-title a") |>
  html_attr("href")

# Keep only articles strictly from /internacional/
# eldiario.es mixes articles from other sections (technology, society...) on its front page
# We identify positions of articles that belong to /internacional/
internacional_idx <- which(str_detect(ref_all, "eldiario\\.es/internacional/"))

# Apply the filter to titles and URLs
titles <- titles_all[internacional_idx]
ref    <- ref_all[internacional_idx]

length(titles)
length(ref)
```

`internacional_idx` is a position vector we will reuse for authors to keep everything aligned.

## Authors

```{r}
# Extract the author of each article
# In eldiario.es the author is in <p class="signature"> inside an <a>
# We extract all authors then filter with internacional_idx
# to maintain alignment with titles and URLs
authors_all <- eldiario_url |>
  html_elements("h2.ni-title") |>
  purrr::map_chr(~ {
    # The author is in the div.ni-info next to the h2
    # We go up to the parent and look for p.signature within the same block
    parent <- xml2::xml_parent(.x)
    a <- parent |>
      html_elements("p.signature a") |>
      html_text2()
    if (length(a) == 0) NA_character_ else paste(a, collapse = ", ")
  })

# Filter with the same indices as titles and URLs
authors <- authors_all[internacional_idx]

length(authors)
authors
```

For each `h2.ni-title` we go up to the parent with `xml_parent()` — which is the `div.md__new` containing both the title and the author — and look for `p.signature` within that same block. This ensures each author matches its title.

## Dates

```{r}
# eldiario.es does not include the date in the URL or in a data-publish attribute
# We set it as NA for now — it will be recovered when scraping each article body
date <- rep(NA_character_, length(titles))
date
```

`rep(NA_character_, length(titles))` creates a vector of NAs — one per article.

## Initial tibble

```{r}
# Build the initial dataset with one row per article
eldiario_df <- tibble(
  title  = titles,
  author = authors,
  ref    = ref,
  date   = date  # NA for now, recovered in the article scraping step
)

eldiario_df
```

## Country list

```{r}
# Build a list of all country names in Spanish from the countrycode package
todos_paises_ed <- countrycode::codelist$cldr.name.es
todos_paises_ed <- todos_paises_ed[!is.na(todos_paises_ed)]

# Build a single regex pattern with all country names
# \b = word boundary — prevents "Irán" from matching inside "Teherán"
patron_paises_ed <- paste0("\\b(", paste(todos_paises_ed, collapse = "|"), ")\\b")
```

## Article scraping function

```{r}
scrape_article_ed <- function(url) {

  Sys.sleep(runif(1, 1, 2))

  tryCatch({
    page <- read_html(url)

    # ── DATE ──────────────────────────────────────────────────────────────────
    fecha <- page |>
      html_elements("time.date") |>
      html_attr("datetime")

    fecha_str <- if (length(fecha) > 0) {
      str_extract(fecha[1], "\\d{4}-\\d{2}-\\d{2}") |>
        as.Date() |>
        format("%d/%m/%Y")
    } else NA_character_

    # ── CITY OF WRITING ORIGIN ─────────────────────────────────────────────
    city <- page |>
      html_elements("span.place") |>
      html_text2() |>
      str_remove("\\s*—\\s*$")

    location <- if (length(city) > 0) city[1] else NA_character_

    # ── COUNTRIES MENTIONED ────────────────────────────────────────────────
    body_text <- page |>
      html_elements("p.article-text") |>
      html_text2() |>
      paste(collapse = " ")

    found <- str_extract_all(body_text, regex(patron_paises_ed, ignore_case = TRUE))[[1]] |>
      unique() |>
      (\(x) todos_paises_ed[str_to_lower(todos_paises_ed) %in% str_to_lower(x)])()

    countries_str <- if (length(found) > 0) paste(found, collapse = ", ") else NA_character_

    # ── POLITICAL FIGURES MENTIONED ────────────────────────────────────────
    figuras_politicas <- c(
      "Pedro Sánchez", "Abascal",
      "Donald Trump", "Von der Leyen",
      "Netanyahu", "Zelenski", "Putin",
      "Milei", "Maduro"
    )

    patron_figuras <- paste0("\\b(", paste(figuras_politicas, collapse = "|"), ")\\b")

    found_figuras <- str_extract_all(
      body_text,
      regex(patron_figuras, ignore_case = TRUE)
    )[[1]] |>
      unique()

    figures_str <- if (length(found_figuras) > 0) {
      paste(found_figuras, collapse = ", ")
    } else NA_character_

    # ── NUMBER OF COMMENTS ─────────────────────────────────────────────────
    # The number is in <p class="comments"> alongside an icon
    # str_extract pulls out only the digits from the text
    comments_text <- page |>
      html_elements("p.comments") |>
      html_text2()

    n_comments <- if (length(comments_text) > 0) {
      str_extract(comments_text[1], "\\d+") |>
        as.integer()
    } else NA_integer_

    # ── RESULT ────────────────────────────────────────────────────────────
    tibble(
      url                 = url,
      date                = fecha_str,
      city_writing_origin = location,
      countries_mentioned = countries_str,
      figures_mentioned   = figures_str,
      n_comments          = n_comments
    )

  }, error = function(e) {
    message("Error in: ", url, " — ", e$message)
    tibble(
      url                 = url,
      date                = NA_character_,
      city_writing_origin = NA_character_,
      countries_mentioned = NA_character_,
      figures_mentioned   = NA_character_,
      n_comments          = NA_integer_
    )
  })
}
```

## Scrape all articles and join

```{r}
article_details_ed <- eldiario_df$ref |>
  map_dfr(scrape_article_ed)

eldiario_df <- eldiario_df |>
  select(-date) |>
  left_join(article_details_ed, by = c("ref" = "url"))

eldiario_df
```

The `left_join` merges the two tibbles by URL — each row in `eldiario_df` is paired with its corresponding row in `article_details_ed`. The `select(-date)` removes the placeholder NA column before joining.

## Clean city of writing origin

```{r}
eldiario_df <- eldiario_df |>
  mutate(
    city_writing_origin = str_remove(city_writing_origin, "(?i)corresponsal en\\s*")
  )

eldiario_df
```

## Clean up environment

```{r}
# Remove all objects except the final dataframe before moving to the next newspaper
rm(list = setdiff(ls(), "eldiario_df"))
gc()
```

------------------------------------------------------------------------

# 2. El Mundo

## Setup

```{r}
library(xml2)
library(rvest)
library(httr)
library(dplyr)
library(stringr)
library(purrr)
library(countrycode)
library(selenium)
library(tibble)

httr::set_config(
  httr::user_agent("Mozilla/5.0 (Windows NT 10.0; Win64; x64); Carlos Morán / carlos.morangq@gmail.com")
)
```

## Download HTML

```{r}
el_mundo_url <- read_html("https://www.elmundo.es/internacional.html")
```

## Titles

```{r}
# Search all <h2> elements with that class and extract clean text
titles <- el_mundo_url |>
  html_elements("h2.ue-c-cover-content__headline") |>
  html_text()

titles
```

## Kicker (extra column)

```{r}
# We iterate article by article because not all articles have a kicker
kicker <- el_mundo_url |>
  html_elements("article.ue-c-cover-content") |>
  purrr::map_chr(~ {
    k <- .x |>
      html_elements("span.ue-c-cover-content__kicker") |>
      html_text2()
    if (length(k) == 0) NA_character_ else k[1]
  })

kicker
```

## URLs

```{r}
ref <- el_mundo_url |>
  html_elements("a.ue-c-cover-content__link") |>
  html_attr("href")

# Keep only URLs containing /internacional/
ref <- ref[grepl("internacional", ref)]

length(ref)
ref
```

Some NAs appear for highlighted editorial articles whose URLs contain tracking parameters (`?intcmp=masnoticiasportada`). These are pinned articles, not organic feed articles.

## Authors

```{r}
# First pass — clean the hidden "Redacción:" prefix added by the screen-reader span
authors <- el_mundo_url |>
  html_elements("article.ue-c-cover-content") |>
  purrr::map_chr(~ {
    nombre <- .x |>
      html_elements("span.ue-c-cover-content__byline-name") |>
      html_text2() |>
      str_remove("^Redacción:\\s*")  # removes "Redacción: " from the start

    rol <- .x |>
      html_elements("span.ue-c-cover-content__byline-role") |>
      html_text2()

    if (length(nombre) == 0) return(NA_character_)

    if (length(rol) > 0) {
      paste(nombre, rol, collapse = ", ")
    } else {
      paste(nombre, collapse = ", ")
    }
  })

authors
```

## Dates

```{r}
# Primary source: data-publish attribute
# Fallback: date embedded in the article URL (YYYY/MM/DD)
date <- el_mundo_url |>
  html_elements("article.ue-c-cover-content") |>
  purrr::map_chr(~ {
    d <- .x |>
      html_elements("[data-publish]") |>
      html_attr("data-publish")

    if (length(d) > 0) return(d[1])

    # If data-publish is missing, try extracting date from the URL
    url <- .x |>
      html_elements("a.ue-c-cover-content__link") |>
      html_attr("href")
    fecha_url <- str_extract(url, "\\d{4}/\\d{2}/\\d{2}")[1]
    if (!is.na(fecha_url)) str_replace_all(fecha_url, "/", "-") else NA_character_
  }) |>
  str_extract("\\d{4}-\\d{2}-\\d{2}") |>
  as.Date(format = "%Y-%m-%d") |>
  format("%d/%m/%Y")

date
```

## Country list and article scraping function

```{r}
all_countries_em <- countrycode::codelist$cldr.name.es
all_countries_em <- all_countries_em[!is.na(all_countries_em)]
# One regex pattern with every country — more efficient than calling each individually
path_countries_em <- paste0("\\b(", paste(all_countries_em, collapse = "|"), ")\\b")

scrape_article_em <- function(ref) {
  Sys.sleep(runif(1, 1, 2))
  tryCatch({
    page <- read_html(ref)
    body <- page |>
      html_elements(xpath = "//div[@class='ue-c-article__body'] | //div[@class='ue-c-article__standfirst']")
    body_text <- body |>
      html_text() |>
      paste(collapse = " ")
    countries_found <- str_extract_all(body_text, path_countries_em)[[1]] |>
      unique() |>
      paste(collapse = ", ")
    if (countries_found == "") return(NA_character_)
    return(countries_found)
  }, error = function(e) {
    message("Error scraping: ", ref, " — ", e$message)
    return(NA_character_)
  })
}
```

## Initial tibble + countries

```{r}
el_mundo_df <- tibble(
  title  = titles,
  kicker = kicker,
  author = authors |> na.omit() |> as.character(),
  ref    = ref,
  date   = date
) |>
  mutate(
    countries_mentioned = purrr::map_chr(ref, scrape_article_em)
  )
```

## City of writing origin

```{r}
city_writing_origin <- el_mundo_url |>
  html_elements("article") |>
  sapply(\(article) {
    span <- article |>
      html_elements("span.ue-c-cover-content__byline-location") |>
      html_text() |>
      str_trim()
    if (length(span) == 0) NA_character_ else span[1]
  })

city_writing_origin

el_mundo_df <- el_mundo_df |>
  mutate(city_writing_origin = city_writing_origin)
```

## Number of comments (Selenium)

```{r}
# Start Selenium server and open Chrome
server <- selenium_server(temp = FALSE)
wait_for_server(server)
driver <- SeleniumSession$new(browser = "chrome")

# Helper function to dismiss the cookie banner if it appears
dismiss_cookies_em <- function(driver) {
  tryCatch({
    cookie_btn <- driver$find_element(
      using = "xpath",
      value = "//button[contains(text(), 'Aceptar y continuar')]"
    )
    cookie_btn$click()
    Sys.sleep(2)
  }, error = function(e) invisible(NULL))
}

# Main function: navigate to each article and extract the comment count
scrape_comments_em <- function(url) {
  Sys.sleep(runif(1, 3, 5))

  tryCatch({
    driver$navigate(url)
    Sys.sleep(5)
    dismiss_cookies_em(driver)
    Sys.sleep(2)

    # Returns text like "85 comentarios" — we extract only the number
    comments <- driver$execute_script("
      var el = document.querySelector('.comments-btn span.ue-c-article__button-text');
      if (el) { return el.innerText; }
      return null;
    ")

    n <- if (is.null(comments) || comments == "") {
      NA_integer_
    } else {
      as.integer(str_extract(comments, "\\d+"))
    }

    message("✅ ", url, " → ", n, " comments")
    tibble(ref = url, n_comments = n)

  }, error = function(e) {
    message("⚠ Error in: ", url, " — ", e$message)
    tryCatch({
      driver <<- SeleniumSession$new(browser = "chrome")
      message("🔄 Session restarted")
    }, error = function(e2) message("❌ Could not restart session"))
    tibble(ref = url, n_comments = NA_integer_)
  })
}

# Run over all URLs
article_comments_em <- el_mundo_df$ref |>
  map_dfr(scrape_comments_em)

# Close the browser
tryCatch(driver$close(), error = function(e) invisible(NULL))
tryCatch(server$stop(),  error = function(e) invisible(NULL))
system("pkill -f chromedriver", ignore.stdout = TRUE, ignore.stderr = TRUE)
system("pkill -f 'chrome --'",  ignore.stdout = TRUE, ignore.stderr = TRUE)
Sys.sleep(5)

# Join with the main tibble
el_mundo_df <- el_mundo_df |>
  left_join(article_comments_em, by = "ref")

el_mundo_df
```

## Political figures mentioned

```{r}
politicians_em <- c(
  "Pedro Sánchez", "Abascal",
  "Donald Trump", "Von der Leyen",
  "Netanyahu", "Zelenski", "Putin",
  "Milei", "Maduro"
)

patron_politicians_em <- paste0("\\b(", paste(politicians_em, collapse = "|"), ")\\b")

scrape_politicians_em <- function(ref) {
  Sys.sleep(runif(1, 1, 2))

  tryCatch({
    page <- read_html(ref)

    body_text <- page |>
      html_elements(xpath = "//div[@class='ue-c-article__body'] | //div[@class='ue-c-article__standfirst']") |>
      html_text(trim = TRUE) |>
      paste(collapse = " ")

    found_politicians <- str_extract_all(
      body_text,
      regex(patron_politicians_em, ignore_case = TRUE)
    )[[1]] |>
      unique()

    politicians_str <- if (length(found_politicians) > 0) {
      paste(found_politicians, collapse = ", ")
    } else NA_character_

    tibble(ref = ref, politicians_mentioned = politicians_str)

  }, error = function(e) {
    message("⚠ Error in: ", ref, " — ", e$message)
    tibble(ref = ref, politicians_mentioned = NA_character_)
  })
}

article_politicians_em <- el_mundo_df$ref |>
  map_dfr(scrape_politicians_em)

el_mundo_df <- el_mundo_df |>
  left_join(article_politicians_em, by = "ref")
```

## Final tibble

```{r}
el_mundo_df
```

## Clean up environment

```{r}
# Remove all objects except the two final dataframes before moving to the next newspaper
rm(list = setdiff(ls(), c("eldiario_df", "el_mundo_df")))
gc()
```

------------------------------------------------------------------------

# 3. El País

## Setup

```{r}
library(xml2)
library(rvest)
library(scrapex)
library(httr)
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(countrycode)
library(selenium)

set_config(
  user_agent("Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/144.0.0.0 Safari/537.36; Gaspar Vigneaux / gasparvigneaux@gmail.com")
)

el_pais_url <- read_html("https://elpais.com/internacional/")
```

## Titles

```{r}
titles <- el_pais_url |>
  html_elements("article header h2 a, article header h3 a") |>
  html_text()

titles
```

## Authors

```{r}
authors <- el_pais_url |>
  html_elements("article") |>
  purrr::map_chr(~ {
    a <- .x |> html_elements("a.c_a_a") |> html_text()
    if (length(a) == 0) NA else paste(a, collapse = ", ")
  })

authors
```

## URLs

```{r}
ref <- el_pais_url |>
  html_elements("article header h2 a, article header h3 a") |>
  xml_attr("href")  # using XPath attribute extraction

ref
```

## Dates

```{r}
# Dates are embedded in the article URL in YYYY-MM-DD format
date <- ref |>
  str_extract("\\d{4}-\\d{2}-\\d{2}") |>
  as.Date(format = "%Y-%m-%d") |>
  format("%d/%m/%Y")
```

## Country list and article scraping function

```{r}
all_countries_ep <- countrycode::codelist$cldr.name.es
all_countries_ep <- all_countries_ep[!is.na(all_countries_ep)]
# One regex pattern with every country — more efficient than calling each individually
path_countries_ep <- paste0("\\b(", paste(all_countries_ep, collapse = "|"), ")\\b")

scrape_article_ep <- function(ref) {
  Sys.sleep(runif(1, 1, 2))
  tryCatch({
    page <- read_html(ref)
    body <- page |>
      html_elements("body div main article header div p, body div main article div.a_c")
    body_text <- body |>
      html_text() |>
      paste(collapse = " ")
    countries_found <- str_extract_all(body_text, path_countries_ep)[[1]] |>
      unique() |>
      paste(collapse = ", ")
    if (countries_found == "") return(NA_character_)
    return(countries_found)
  }, error = function(e) {
    message("Error scraping: ", ref, " — ", e$message)
    return(NA_character_)
  })
}
```

## Political figures scraping function

```{r}
scrape_figures_ep <- function(ref) {
  Sys.sleep(runif(1, 1, 2))
  tryCatch({
    page <- read_html(ref)
    body_text <- page |>
      html_elements("body div main article header div p, body div main article div.a_c") |>
      html_text() |>
      paste(collapse = " ")

    figuras_politicas <- c(
      "Pedro Sánchez", "Abascal",
      "Donald Trump", "Von der Leyen",
      "Netanyahu", "Zelenski", "Putin",
      "Milei", "Maduro"
    )
    patron_figuras <- paste0("\\b(", paste(figuras_politicas, collapse = "|"), ")\\b")
    found <- str_extract_all(body_text, regex(patron_figuras, ignore_case = TRUE))[[1]] |> unique()
    if (length(found) > 0) paste(found, collapse = ", ") else NA_character_

  }, error = function(e) NA_character_)
}
```

## Initial tibble + countries + political figures

```{r}
el_pais_df <- tibble(
  title  = titles,
  author = authors,
  ref    = ref,
  date   = date
) |>
  mutate(
    countries_mentioned = purrr::map_chr(ref, scrape_article_ep),
    figures_mentioned   = purrr::map_chr(ref, scrape_figures_ep)
  )
```

## City of writing origin

```{r}
city_writing_origin <- el_pais_url |>
  html_elements("article") |>
  sapply(\(article) {
    span <- article |>
      html_elements("span.c_a_l") |>
      html_text() |>
      str_trim()
    if (length(span) == 0) NA_character_ else span[1]
  })

city_writing_origin

el_pais_df <- el_pais_df |>
  mutate(city_writing_origin = city_writing_origin)
```

## Number of comments (Selenium)

```{r}
# Start session
server <- selenium_server(temp = FALSE)
wait_for_server(server)
driver <- SeleniumSession$new(browser = "chrome")

# Accept cookies once on the first URL
driver$navigate(el_pais_df$ref[1])
Sys.sleep(4)
tryCatch({
  cookie_btn <- driver$find_element(
    using = "xpath",
    value = "//button[contains(text(), 'Aceptar y continuar') or contains(text(), 'Aceptar')]"
  )
  cookie_btn$click()
  message("✅ Cookie banner dismissed")
  Sys.sleep(4)
}, error = function(e) {
  message("⚠ No cookie banner found")
})

scrape_comments_ep <- function(ref) {
  Sys.sleep(runif(1, 3, 5))

  tryCatch({
    driver$navigate(ref)
    Sys.sleep(5)

    comments <- driver$execute_script("
      var el = document.querySelector('span.disqus-comment-count');
      if (el) { return el.innerText; }
      else { return null; }
    ")

    n <- if (is.null(comments) || comments == "") NA_integer_ else as.integer(comments)
    message("✅ ", ref, " → ", n, " comments")
    tibble(ref = ref, n_comments = n)

  }, error = function(e) {
    message("⚠ Error in: ", ref, " — ", e$message)
    tibble(ref = ref, n_comments = NA_integer_)
  })
}

# Run over all articles
article_comments_ep <- el_pais_df$ref |>
  map_dfr(scrape_comments_ep)

# Join with main tibble
el_pais_df <- el_pais_df |>
  left_join(article_comments_ep, by = "ref") |>
  rename(n_comments = n_comments)

tryCatch(driver$close(), error = function(e) invisible(NULL))
tryCatch(server$stop(),  error = function(e) invisible(NULL))
system("pkill -f chromedriver", ignore.stdout = TRUE, ignore.stderr = TRUE)
system("pkill -f 'chrome --'",  ignore.stdout = TRUE, ignore.stderr = TRUE)
Sys.sleep(5)

el_pais_df
```

## Final tibble

```{r}
el_pais_df
```

## Clean up environment

```{r}
# Remove all objects except the three final dataframes before moving to the next newspaper
rm(list = setdiff(ls(), c("eldiario_df", "el_mundo_df", "el_pais_df")))
gc()
```

------------------------------------------------------------------------

# 4. La Vanguardia

## Setup

```{r}
library(scrapex)
library(rvest)
library(httr)
library(dplyr)
library(tidyr)
library(stringr)
library(purrr)
library(xml2)
library(countrycode)
library(selenium)

set_config(
  user_agent("Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/134.0.0.0 Safari/537.36")
)

la_vanguardia_url <- read_html("https://www.lavanguardia.com/internacional")
```

## Titles

```{r}
titles <- la_vanguardia_url |>
  html_elements("article div h2 a") |>
  html_text()

titles
```

## Authors

```{r}
authors_lv <- la_vanguardia_url |>
  html_elements("article div div.author, article div div span") |>
  html_text()

authors_lv
```

## URLs

```{r}
url_news_lv <- la_vanguardia_url |>
  html_elements("article div h2 a") |>
  xml_attr("href")

url_news_lv
```

## Dates

```{r}
# Dates are embedded in the URL as an 8-digit string (YYYYMMDD)
dates_lv <- url_news_lv |>
  str_extract("\\d{8}") |>
  as.Date(format = "%Y%m%d") |>
  format("%d/%m/%Y")

dates_lv
```

## Initial tibble

```{r}
la_vanguardia_df <- tibble(
  title  = str_squish(titles),
  author = str_squish(authors_lv),
  url    = paste0("https://www.lavanguardia.com", url_news_lv),
  date   = dates_lv
)

la_vanguardia_df
```

## City of writing origin

```{r}
scrape_city_lv <- function(url) {
  Sys.sleep(runif(1, 1, 2))

  tryCatch({
    page <- read_html(url)

    location <- page |>
      html_element("p.author_location span") |>
      html_text(trim = TRUE) |>
      str_extract("^[^.]+") |>
      str_squish()

    if (is.na(location) || location == "") location <- NA_character_

    tibble(url = url, city_writing_origin = location)

  }, error = function(e) {
    message("⚠ Error in: ", url, " — ", e$message)
    tibble(url = url, city_writing_origin = NA_character_)
  })
}

# Iterate over all articles
article_cities_lv <- la_vanguardia_df$url |>
  map_dfr(scrape_city_lv)

# Join with the main tibble
la_vanguardia_df <- la_vanguardia_df |>
  left_join(article_cities_lv, by = "url")

la_vanguardia_df
```

## Countries mentioned

```{r}
# Build the country list BEFORE the function
todos_paises_lv <- countrycode::codelist$cldr.name.es
todos_paises_lv <- todos_paises_lv[!is.na(todos_paises_lv)]
patron_paises_lv <- paste0("\\b(", paste(todos_paises_lv, collapse = "|"), ")\\b")

scrape_countries_lv <- function(url) {
  Sys.sleep(runif(1, 1, 2))

  tryCatch({
    page <- read_html(url)

    body_text <- page |>
      html_elements("div.article_story") |>
      html_text(trim = TRUE) |>
      paste(collapse = " ")

    found <- str_extract_all(body_text, regex(patron_paises_lv, ignore_case = TRUE))[[1]] |>
      unique() |>
      (\(x) todos_paises_lv[str_to_lower(todos_paises_lv) %in% str_to_lower(x)])()

    countries_str <- if (length(found) > 0) paste(found, collapse = ", ") else NA_character_

    tibble(url = url, countries_mentioned = countries_str)

  }, error = function(e) {
    message("⚠ Error in: ", url, " — ", e$message)
    tibble(url = url, countries_mentioned = NA_character_)
  })
}

# Iterate over all articles
article_countries_lv <- la_vanguardia_df$url |>
  map_dfr(scrape_countries_lv)

# Join with the main tibble
la_vanguardia_df <- la_vanguardia_df |>
  left_join(article_countries_lv, by = "url")

la_vanguardia_df
```

## Number of comments (Selenium)

```{r}
# Start session
server <- selenium_server(temp = FALSE)
wait_for_server(server)
driver <- SeleniumSession$new(browser = "chrome")

driver$navigate(la_vanguardia_df$url[2])
Sys.sleep(4)

# Accept cookies
tryCatch({
  cookie_btn <- driver$find_element(
    using = "xpath",
    value = "//button[contains(text(), 'Aceptar y continuar')]"
  )
  cookie_btn$click()
  message("✅ Cookie banner dismissed")
  Sys.sleep(4)
}, error = function(e) {
  message("⚠ No cookie banner found")
})

# Verify comment element
comments_test <- driver$execute_script("
  var el = document.querySelector('#social-comments-count');
  if (el) { return el.innerText; } else { return 'element not found'; }
")
cat("Comments found:", comments_test, "\n")

scrape_comments_lv <- function(url) {
  Sys.sleep(runif(1, 3, 5))

  tryCatch({
    driver$navigate(url)
    Sys.sleep(4)

    comments <- driver$execute_script("
      var el = document.querySelector('#social-comments-count');
      if (el) { return el.innerText; }
      else { return null; }
    ")

    tibble(url = url, n_comments = ifelse(is.null(comments), NA_character_, comments))

  }, error = function(e) {
    message("⚠ Error in: ", url, " — ", e$message)
    tibble(url = url, n_comments = NA_character_)
  })
}

# Run over all articles
article_comments_lv <- la_vanguardia_df$url |>
  map_dfr(scrape_comments_lv)

# Join with main tibble
la_vanguardia_df <- la_vanguardia_df |>
  left_join(article_comments_lv, by = "url")

la_vanguardia_df
```

## Political figures mentioned

```{r}
politicians_lv <- c(
  "Pedro Sánchez", "Abascal",
  "Donald Trump", "Von der Leyen",
  "Netanyahu", "Zelenski", "Putin",
  "Milei", "Maduro"
)

patron_politicians_lv <- paste0("\\b(", paste(politicians_lv, collapse = "|"), ")\\b")

scrape_politicians_lv <- function(url) {
  Sys.sleep(runif(1, 1, 2))

  tryCatch({
    page <- read_html(url)

    body_text <- page |>
      html_elements("div.article_story") |>
      html_text(trim = TRUE) |>
      paste(collapse = " ")

    found_politicians <- str_extract_all(
      body_text,
      regex(patron_politicians_lv, ignore_case = TRUE)
    )[[1]] |>
      unique()

    politicians_str <- if (length(found_politicians) > 0) {
      paste(found_politicians, collapse = ", ")
    } else NA_character_

    tibble(url = url, politicians_mentioned = politicians_str)

  }, error = function(e) {
    message("⚠ Error in: ", url, " — ", e$message)
    tibble(url = url, politicians_mentioned = NA_character_)
  })
}

# Iterate over all articles
article_politicians_lv <- la_vanguardia_df$url |>
  map_dfr(scrape_politicians_lv)

# Join with main tibble
la_vanguardia_df <- la_vanguardia_df |>
  left_join(article_politicians_lv, by = "url")

la_vanguardia_df
```

## Clean up environment

```{r}
# Remove all objects except the four final dataframes
rm(list = setdiff(ls(), c("eldiario_df", "el_mundo_df", "el_pais_df", "la_vanguardia_df")))
gc()
```

------------------------------------------------------------------------

# 5. Merge all datasets

We standardize column names across the four datasets, add `newspaper` and `article_id` columns, and export a single CSV.

```{r}
library(dplyr)
library(readr)

# ── Standardize column names ───────────────────────────────────────────────────

# El Diario: rename ref → url, figures_mentioned → politicians_mentioned
eldiario_clean <- eldiario_df |>
  rename(
    url                  = ref,
    politicians_mentioned = figures_mentioned
  ) |>
  mutate(newspaper = "eldiario.es")

# El Mundo: rename ref → url; drop kicker column
el_mundo_clean <- el_mundo_df |>
  rename(url = ref) |>
  select(-any_of("kicker")) |>
  mutate(newspaper = "elmundo.es")

# El País: rename ref → url, countries → countries_mentioned, figures_mentioned stays
el_pais_clean <- el_pais_df |>
  rename(url = ref) |>
  rename(any_of(c(countries_mentioned = "countries",
                  politicians_mentioned = "figures_mentioned"))) |>
  mutate(newspaper = "elpais.com")

# La Vanguardia: url column already named url
la_vanguardia_clean <- la_vanguardia_df |>
  mutate(
    n_comments = as.integer(str_extract(as.character(n_comments), "\\d+")),
    newspaper = "lavanguardia.com"
  )

# ── Bind all rows ──────────────────────────────────────────────────────────────

all_newspapers <- bind_rows(
  eldiario_clean,
  el_mundo_clean,
  el_pais_clean,
  la_vanguardia_clean
)

# ── Add article_id ─────────────────────────────────────────────────────────────
# A unique sequential ID per article across all newspapers
all_newspapers <- all_newspapers |>
  mutate(article_id = row_number()) |>
  select(article_id, newspaper, everything())

all_newspapers
```

## Export to CSV

```{r}
write_csv(
  all_newspapers,
  file = paste0("all_newspapers_", format(Sys.Date(), "%d_%m_%Y"), ".csv")
)

message("✅ CSV exported: all_newspapers_", format(Sys.Date(), "%d_%m_%Y"), ".csv")
```

## EDA

```{r}
library(DataExplorer)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(forcats)

plot_intro(all_newspapers)
introduce(all_newspapers)
```


## Graphs and Tables

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(forcats)

# Shared colour palette — one colour per newspaper
paper_colors <- c(
  "eldiario.es"      = "#E63946",
  "elmundo.es"       = "#457B9D",
  "elpais.com"       = "#2D6A4F",
  "lavanguardia.com" = "#E9C46A"
)

theme_papers <- function() {
  theme_minimal(base_size = 13) +
    theme(
      plot.title       = element_text(face = "bold", size = 15, margin = margin(b = 6)),
      plot.subtitle    = element_text(colour = "grey45", size = 11, margin = margin(b = 10)),
      axis.title       = element_text(size = 11),
      legend.position  = "bottom",
      legend.title     = element_blank(),
      panel.grid.minor = element_blank()
    )
}

# Total articles per newspaper — denominator used throughout
totals <- all_newspapers |>
  count(newspaper, name = "total_articles")
```


### Articles per newspaper

```{r articles-table}
articles_per_paper <- totals |>
  mutate(pct = round(total_articles / sum(total_articles) * 100, 1))

articles_per_paper
```

```{r articles-plot}
ggplot(articles_per_paper,
       aes(x = fct_reorder(newspaper, pct), y = pct, fill = newspaper)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_text(aes(label = paste0(pct, "%")), hjust = -0.3,
            size = 4.5, fontface = "bold") +
  scale_fill_manual(values = paper_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  coord_flip() +
  labs(
    title    = "Share of total articles by newspaper",
    subtitle = "% of all scraped articles",
    x = NULL, y = "% of corpus"
  ) +
  theme_papers()
```

------------------------------------------------------------------------

### Average comments per newspaper

```{r avg-comments-table}
avg_comments <- all_newspapers |>
  filter(!is.na(n_comments)) |>
  group_by(newspaper) |>
  summarise(
    avg_comments    = round(mean(n_comments), 1),
    median_comments = round(median(n_comments), 1),
    n_articles      = n(),
    .groups = "drop"
  )

avg_comments
```

```{r avg-comments-plot}
ggplot(avg_comments,
       aes(x = fct_reorder(newspaper, avg_comments), y = avg_comments,
           fill = newspaper)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_text(aes(label = avg_comments), hjust = -0.3,
            size = 4.5, fontface = "bold") +
  scale_fill_manual(values = paper_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  coord_flip() +
  labs(
    title    = "Average comments per article by newspaper",
    subtitle = "Mean comments per published article (NA excluded)",
    x = NULL, y = "Avg. comments per article"
  ) +
  theme_papers()
```


### Top 10 most commented articles

Comments are shown as a ratio to each newspaper's own average, so a score of `3x` means the article received 3 times more comments than the typical article from that outlet. The dashed line marks `1x` (the newspaper average).

```{r top10-comments-table}
newspaper_avg <- all_newspapers |>
  filter(!is.na(n_comments)) |>
  group_by(newspaper) |>
  summarise(avg = mean(n_comments), .groups = "drop")

top10_comments <- all_newspapers |>
  filter(!is.na(n_comments)) |>
  left_join(newspaper_avg, by = "newspaper") |>
  mutate(comments_ratio = round(n_comments / avg, 1)) |>
  slice_max(order_by = comments_ratio, n = 10) |>
  mutate(
    short_title = str_trunc(title, 52),
    short_title = fct_reorder(short_title, comments_ratio)
  )

top10_comments |>
  select(newspaper, title, n_comments, comments_ratio) |>
  arrange(desc(comments_ratio))
```



### Most frequently mentioned countries

```{r countries-table}
countries_long <- all_newspapers |>
  filter(!is.na(countries_mentioned)) |>
  select(article_id, newspaper, countries_mentioned) |>
  mutate(country = str_split(countries_mentioned, ",\\s*")) |>
  unnest(country) |>
  mutate(country = str_squish(country)) |>
  filter(country != "")

top10_countries <- countries_long |>
  count(country, sort = TRUE) |>
  slice_head(n = 10) |>
  pull(country)

country_rates <- countries_long |>
  filter(country %in% top10_countries) |>
  count(country, newspaper) |>
  left_join(totals, by = "newspaper") |>
  mutate(pct = round(n / total_articles * 100, 1)) |>
  group_by(country) |>
  mutate(avg_pct = mean(pct)) |>
  ungroup() |>
  mutate(country = fct_reorder(country, avg_pct))

country_rates |>
  select(country, newspaper, n, pct) |>
  arrange(desc(pct))
```

```{r countries-plot}
ggplot(country_rates,
       aes(x = country, y = pct, fill = newspaper)) +
  geom_col(position = "dodge", width = 0.7) +
  scale_fill_manual(values = paper_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.08)),
                     labels = function(x) paste0(x, "%")) +
  coord_flip() +
  labs(
    title    = "Top 10 countries mentioned — share of each newspaper's articles",
    subtitle = "% of articles per newspaper that mention each country",
    x = NULL, y = "% of newspaper's articles"
  ) +
  theme_papers()
```



### Most frequently mentioned politicians

Each cell shows the percentage of that newspaper's articles that mention the political figure. Empty cells mean zero mentions.

```{r politicians-table}
politicians_long <- all_newspapers |>
  filter(!is.na(politicians_mentioned)) |>
  select(article_id, newspaper, politicians_mentioned) |>
  mutate(politician = str_split(politicians_mentioned, ",\\s*")) |>
  unnest(politician) |>
  mutate(politician = str_squish(politician)) |>
  filter(politician != "")

heatmap_data <- politicians_long |>
  count(politician, newspaper) |>
  left_join(totals, by = "newspaper") |>
  mutate(pct = round(n / total_articles * 100, 1)) |>
  complete(politician, newspaper, fill = list(pct = 0)) |>
  group_by(politician) |>
  mutate(avg_pct = mean(pct)) |>
  ungroup() |>
  mutate(politician = fct_reorder(politician, avg_pct))

heatmap_data |>
  select(politician, newspaper, pct) |>
  arrange(desc(pct))
```

```{r politicians-plot}
ggplot(heatmap_data,
       aes(x = newspaper, y = politician, fill = pct)) +
  geom_tile(colour = "white", linewidth = 0.6) +
  geom_text(aes(label = ifelse(pct > 0, paste0(pct, "%"), "")),
            size = 3.8, fontface = "bold", colour = "white") +
  scale_fill_gradient(low = "#d4e6f1", high = "#1a5276",
                      name = "% of articles") +
  scale_x_discrete(guide = guide_axis(angle = 20)) +
  labs(
    title    = "Political figures — share of each newspaper's articles",
    subtitle = "% of articles per newspaper that mention each figure",
    x = NULL, y = NULL
  ) +
  theme_papers() +
  theme(panel.grid = element_blank(), legend.position = "right")
```


### Most usual authors by newspaper

```{r authors-table}
top_authors <- all_newspapers |>
  filter(!is.na(author), author != "") |>
  count(newspaper, author) |>
  left_join(totals, by = "newspaper") |>
  mutate(pct = round(n / total_articles * 100, 1)) |>
  group_by(newspaper) |>
  slice_max(order_by = pct, n = 5, with_ties = FALSE) |>
  ungroup() |>
  mutate(author = fct_reorder(author, pct))

top_authors |>
  select(newspaper, author, n, pct) |>
  arrange(desc(n))
```




### Cities of writing origin

Only cities that account for at least 2% of any single newspaper's articles are shown, to keep the chart readable.

```{r cities-table}
cities_long <- all_newspapers |>
  filter(!is.na(city_writing_origin), city_writing_origin != "") |>
  mutate(city = str_squish(city_writing_origin)) |>
  count(city, newspaper) |>
  left_join(totals, by = "newspaper") |>
  mutate(pct = round(n / total_articles * 100, 1))

top_cities <- cities_long |>
  group_by(city) |>
  filter(any(pct >= 2)) |>
  mutate(avg_pct = mean(pct)) |>
  ungroup() |>
  mutate(city = fct_reorder(city, avg_pct))

top_cities |>
  group_by(city) |>
  summarise(
    avg_pct    = round(mean(pct), 1),
    newspapers = paste(newspaper, collapse = ", "),
    .groups    = "drop"
  ) |>
  arrange(desc(avg_pct))
```

```{r cities-plot}
ggplot(top_cities,
       aes(x = city, y = pct, fill = newspaper)) +
  geom_col(position = "dodge", width = 0.7) +
  scale_fill_manual(values = paper_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.08)),
                     labels = function(x) paste0(x, "%")) +
  coord_flip() +
  labs(
    title    = "Cities of writing origin",
    subtitle = "% of each newspaper's articles filed from each city",
    x = NULL, y = "% of newspaper's articles"
  ) +
  theme_papers()
```

